### 2023-04-11 main.tex
organization

Consider the Lie algebra $gl(N)$ with generators denoted by $\EE_{ab}$ with $a,b\in \{1,\ldots,N\}$ and commutation relations
\begin{equation}\label{eq:comgl}
\left[\EE_{ab},\EE_{cd}\right]=\EE_{ad}\delta_{bc}-\EE_{cb}\delta_{ad}\qquad \forall a,b\in \{1,\ldots,N\}
\end{equation}
The finite-dimensional representations are labelled by partitions $\lambda=(\lambda_1,\lambda_2,\ldots,\lambda_N)$ of $\nu$ with $\lambda_i\in \mathbb{N}$ and $\sum_{i=1}^N \lambda_i = \nu$. We assume $\lambda_i\geq \lambda_{i+1}$ without loss of generality.
We are interested in the {\em symmetric} finite-dimensional representations with 
\begin{equation}\label{eq:dynkin}
    \lambda=(\twoj,0,\ldots,0) \qquad\text{where}\qquad \twoj\in\mathbb{N}
\end{equation} 
The dimension $M$ of the symmetric representations is given by the combination of $N$ objects in $\twoj$ positions with repetition, namely
\begin{equation}
	M= \frac{(N+\twoj-1)!}{\twoj  !(N-1)!}
\end{equation} 
The basis elements of the vector space $\mathbb{C}^{M}$ are the column vectors 
\begin{equation}
  |n\rangle=  |n_{1},\ldots,n_{N}\rangle,\quad \text{with}\quad n_{i}\in\mathbb{N}_{0}\quad \text{sucht that}\quad \sum_{i=1}^{N}n_{i}=\nu
\end{equation}
%where $\sum_{a=1}^{N}n_{i}=\nu$. 
This basis vector satisfy the orthogonality relation with respect to the Euclidean scalar product 
\begin{equation}\label{ortho}
    \langle m_{1},\ldots,m_{N}|n_{1},\ldots,n_{N}\rangle=\prod_{a=1}^{N}\delta_{m_{a},n_{a}}
\end{equation}
where  $ \langle m_{1},\ldots,m_{N}|$ is the row vector obtained by transposing $|m_{1},\ldots,m_{N}\rangle$ and $\delta_{m_{a},n_{a}}$ is the Kronhecker delta. 
The explicit actions of the algebra generators on the vectors $|n\rangle$ are the following:
\begin{equation}\label{actionE}
	\begin{cases}
		E_{ab}|n_{1},\ldots,n_{a},\ldots,n_{b},\ldots,n_{N}\rangle =n_{b}|n_{1},\ldots,n_{a}+1,\ldots,n_{b}-1,\ldots,n_{N}\rangle\quad a\neq b\\[0.1cm]
		E_{aa}|n_{1},\ldots,n_{a},\ldots,n_{b},\ldots,n_{N}\rangle = n_{a} |n_{1},\ldots,n_{a},\ldots,n_{b},\ldots,n_{N}\rangle\quad a=b
	\end{cases}
\end{equation}  
One can check that the matrices defined in this way satisfy the commutation relations \eqref{eq:comgl} and yield Dynkin weight \eqref{eq:dynkin}.

\noindent
\textbf{Reamrk}: The symmetric representations $\lambda=(\twoj,0,\ldots,0)$ are dual to the representations $\lambda=(\twoj,\ldots,\twoj,0)$ that can be obtained via 
\begin{equation}
   \bar E_{ab}=\nu\delta_{ab}-E_{N-b+1,N-a+1}
\end{equation}
{\color{red} This needs to be checked and give alternative form of Hamiltonian?}\\
By the Lie algebra above we describe the process with generator \eqref{Generator}. The state space \eqref{stateSpace} is the vector space with basis elements

@@

\begin{equation}\label{transpositionPropertyH}
    H_{x,y}^{T}=\left(R_{x}R_{y}\right)H_{x,y}\left(R_{x}R_{y}\right)^{-1}
\end{equation}
therefore we have found the symilarity tranfsormation between $H_{x,y}$ and its transposed. \\
As already pointed out, $H_{x,y}$ is a linear function of $\Delta(C_{2})$. This implies that symmetries of  $\Delta(C_{2})$ are also symmetries of  $H_{x,y}$. Moreover, the coproduct is a Lie algebra homomorphism, i.e. $\forall A,B\in gl(N)$
\begin{equation}\label{symmetryCoproduct}
    [A,B]=0\quad \Longrightarrow\quad \left[\Delta(A),\Delta(B)\right]=0
\end{equation}
This tells us that it is enough to find a symmetry of $C_{2}$. This element belongs to the center of $gl(N)$, and therefore it commutes with all the element of this algebra. Consider
\begin{equation}\label{equationE}
    E=\sum_{a=2}^{N}E_{a1}
\end{equation}
To obtain a product structure of the elements of the duality matrix, we introduce
\begin{equation}
    S=\exp{(\Delta(E))}
\end{equation}
Finally, for fixed $(x,y)\in \mathcal{E}$, we define
\begin{equation}
    S_{x,y}=\exp{(\Delta^{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta^{x,y}(E)$ is a copy of the coproduct acting on sites $x$ and $y$. This operator $S^{x,y}$ satisfies
\begin{equation}\label{symmetryH}
    \left[S_{x,y},H_{x,y}\right]=0
\end{equation}
i.e. $S_{x,y}$ is a symmetry of $H_{x,y}$. \\ Exploiting these considerations we write
\begin{equation}
    \begin{split}
        H_{x,y}^{T}D&=(R_{x}R_{y})H_{x,y}(R_{x}R_{y})^{-1}\left(d_{x}\otimes\dd\right)\left(d_{y}\otimes\mathcal{D}_{\widehat{y}}\right)\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{\widehat{z}}\right)
        \\&=\left(R_{x}\exp{(E^{x})}\otimes \dd\right)\left(R_{y}\exp{(E^{y})}\otimes \mathcal{D}_{\widehat{y}}\right)H_{x,y}\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{\widehat{z}}\right)
        \\&=
        DH_{x,y}
    \end{split}
\end{equation}
 where we used \eqref{transpositionPropertyH} and \eqref{symmetryH} in the second equality. Thus, \eqref{edgeDualRealtion} is proved. 
 %\textbf{Reamrk}: it is important to notice that the elements of the duality matrix \eqref{dualityElements} are well defined only if at each site $x\in V$ and for every species $a\in \{2,\ldots,N\}$ the number of dual particles $r_{x}^{a}$ is lower or equal than the number of original particles $n_{a}^{x}$. This implies that the dynamics of the dual process is simpler, because described by a lower number of particles. By the way, it is possible to perform computations similar to the one made in this proof choosing instead of  \eqref{equationE} {\color{red} $E=\sum_{a=2}^{N}E_{1a}$ or $E=\sum_{a=2}^{N}E_{aa}$.}
 %{\color{blue} That's confusing. What is E?}

@@

@@

\subsection{Linear reaction-diffusion process on a finite graph}
\paragraph{Action on the occupation variable of the edge generator}
\paragraph{Action on the occupation variable of the site generator}
\subsection{Reaction-diffusion process on a chain}
\subsubsection{Proof of duality relation for reaction-diffusion process}
With this prove we answered the second and third questions stated at the beginning of this section.

@@

\bibliography{ref}
\bibliographystyle{unsrt}
\end{document}

\appendix
\section{Oscillator}

\textbf{Reamrk}: The basis element the we have defined above, can also be thought as proper compositions of the rising and lowering operators of th Heisenberg Lie algebra. Indeed, \begin{equation}\label{creationOperators}

@@

The basis elements of $gl(N)$ can be written as
		e_{ab}=\oad_{a}\oa_{b}\qquad \forall a,b=1,\ldots N

@@

\begin{equation}
	\begin{split}
		H=\sum_{x,y\in \mathcal{E}}\omega_{x,y}H_{x,y}+\sum_{x\in V}\Gamma_{x}H_{x}
	\end{split}
\end{equation}
where the edge Hamiltonian is
\begin{equation}
H_{x,y}=\sum_{a,l=1}^{N}\left(e_{al}\otimes e_{la}-e_{ll}\otimes e_{aa}\right)
 \end{equation}
 and where the site Hamiltonian is
 \begin{equation}
H_{x}=\sum_{a,l=1}^{N}\alpha_{a}^{x}\left(e_{a,l}-e_{ll}\right)
\end{equation}
Moreover, by the \eqref{coproductC2} and by computing the explicit action of the diagonal part, we can write this Hamiltonian in function of the coproduct of the second Casimir
\begin{equation}
	H=\sum_{x,y\in \mathcal{E}}\omega_{x,y}\left\{\frac{1}{2}\Delta(C_{2})-2j(4j+N)\frac{1}{2}\mathbbm{1}\otimes\mathbbm{1}\right\}+\sum_{x\in V}\Gamma_{x}\sum_{a,l=1}^{N}\alpha_{a}^{x}\left(e_{a,l}-2j\mathbbm{1}\right)
\end{equation}

>>>>

Consider the Lie algebra $gl(N)$ with generators denoted by $\EE_{ab}$ with $a,b\in \{1,\ldots,N\}$ and commutation relations
\begin{equation}\label{eq:comgl}
\left[\EE_{ab},\EE_{cd}\right]=\EE_{ad}\delta_{bc}-\EE_{cb}\delta_{ad}\qquad \forall a,b\in \{1,\ldots,N\}
\end{equation}
The finite-dimensional representations are labelled by partitions $\lambda=(\lambda_1,\lambda_2,\ldots,\lambda_N)$ of $\nu$ with $\lambda_i\in \mathbb{N}$ and $\sum_{i=1}^N \lambda_i = \nu$. We assume $\lambda_i\geq \lambda_{i+1}$ without loss of generality.
We are interested in the {\em symmetric} finite-dimensional representations with 
\begin{equation}\label{eq:dynkin}
    \lambda=(\twoj,0,\ldots,0) \qquad\text{where}\qquad \twoj\in\mathbb{N}
\end{equation} 
The dimension $M$ of the symmetric representations is given by the combination of $N$ objects in $\twoj$ positions with repetition, namely
\begin{equation}
	M= \frac{(N+\twoj-1)!}{\twoj  !(N-1)!}
\end{equation} 
The basis elements of the vector space $\mathbb{C}^{M}$ are the column vectors 
\begin{equation}
  |n\rangle=  |n_{1},\ldots,n_{N}\rangle,\quad \text{with}\quad n_{i}\in\mathbb{N}_{0}\quad \text{sucht that}\quad \sum_{i=1}^{N}n_{i}=\nu
\end{equation}
%where $\sum_{a=1}^{N}n_{i}=\nu$. 
This basis vector satisfy the orthogonality relation with respect to the Euclidean scalar product 
\begin{equation}\label{ortho}
    \langle m_{1},\ldots,m_{N}|n_{1},\ldots,n_{N}\rangle=\prod_{a=1}^{N}\delta_{m_{a},n_{a}}
\end{equation}
where  $ \langle m_{1},\ldots,m_{N}|$ is the row vector obtained by transposing $|m_{1},\ldots,m_{N}\rangle$ and $\delta_{m_{a},n_{a}}$ is the Kronhecker delta. 
The explicit actions of the algebra generators on the vectors $|n\rangle$ are the following:
\begin{equation}\label{actionE}
	\begin{cases}
		E_{ab}|n_{1},\ldots,n_{a},\ldots,n_{b},\ldots,n_{N}\rangle =n_{b}|n_{1},\ldots,n_{a}+1,\ldots,n_{b}-1,\ldots,n_{N}\rangle\quad a\neq b\\[0.1cm]
		E_{aa}|n_{1},\ldots,n_{a},\ldots,n_{b},\ldots,n_{N}\rangle = n_{a} |n_{1},\ldots,n_{a},\ldots,n_{b},\ldots,n_{N}\rangle\quad a=b
	\end{cases}
\end{equation}  
One can check that the matrices defined in this way satisfy the commutation relations \eqref{eq:comgl} and yield Dynkin weight \eqref{eq:dynkin}. \\
\newline 
\textbf{Remark}: it is possible to define these matrices in terms of creation and annihilation operators: see Appendix \ref{appA}.
\\ \\
\textbf{Reamrk}: The symmetric representations $\lambda=(\twoj,0,\ldots,0)$ are dual to the representations $\lambda=(\twoj,\ldots,\twoj,0)$ that can be obtained via 
\begin{equation}
   \bar E_{ab}=\nu\delta_{ab}-E_{N-b+1,N-a+1}
\end{equation}
{\color{red} This needs to be checked and give alternative form of Hamiltonian?}\\
By the Lie algebra above we describe the process with generator \eqref{Generator}. The state space \eqref{stateSpace} is the vector space with basis elements

@@

\begin{equation}\label{transpositionPropertyH}
    H_{x,y}^{T}=\left(R_{x}R_{y}\right)H_{x,y}\left(R_{x}R_{y}\right)^{-1}
\end{equation}
therefore we have found the symilarity tranfsormation between $H_{x,y}$ and its transposed. \\
We now look for a symmetry of $H_{x,y}$, i.e. a matrix $S_{x,y}$ of the same dimension such that it satisfies 
\begin{equation}
	H_{x,y}S_{x,y}=S_{s,y}H_{x,y}\quad \Leftrightarrow \quad [H_{x,y},S_{x,y}]=0
\end{equation}
Using \eqref{hamiltonianCasimir}, we observe that $H_{x,y}$ is proportional to the coproduct of the second Casimir, up to a diagonal term. Therefore, it is enough to look for a symmetry of $\Delta (C_{2})$. Using the bilinearity of the coproduct operator and the fact that $C_{2}$ belongs to the centre of the Lie algebra, it is easy to show that for any matrix $E\in gl(N)$
\begin{equation}
	[C_{2},E]=0\quad \Rightarrow\quad \left[\Delta (C_{2},\Delta(A)) \right]=0
\end{equation}
We chose 
\begin{equation}
	E=\sum_{a=2}^{N}E_{a1}
\end{equation}
and, for a fixed $(x,y)\in \mathcal{E}$, we define
\begin{equation}
	S_{x,y}=\exp{(\Delta^{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta^{x,y}(E)$ is a copy of the coproduct acting on sites $x$ and $y$. Therefore, this operator $S_{x,y}$ satisfies
\begin{equation}\label{symmetryH}
	\left[S_{x,y},H_{x,y}\right]=0
\end{equation}
i.e. it is a symmetry of $H_{x,y}$. \\ Exploiting these considerations we write

\begin{comment}
As already pointed out, $H_{x,y}$ is a linear function of $\Delta(C_{2})$. This implies that symmetries of  $\Delta(C_{2})$ are also symmetries of  $H_{x,y}$. Moreover, the coproduct is a Lie algebra homomorphism, i.e. $\forall A,B\in gl(N)$
\begin{equation}\label{symmetryCoproduct}
    [A,B]=0\quad \Longrightarrow\quad \left[\Delta(A),\Delta(B)\right]=0
\end{equation}
This tells us that it is enough to find a symmetry of $C_{2}$. This element belongs to the center of $gl(N)$, and therefore it commutes with all the element of this algebra. Consider
\begin{equation}\label{equationE}
    E=\sum_{a=2}^{N}E_{a1}
\end{equation}
To obtain a product structure of the elements of the duality matrix, we introduce
\begin{equation}
    S=\exp{(\Delta(E))}
\end{equation}
Finally, for fixed $(x,y)\in \mathcal{E}$, we define
\begin{equation}
    S_{x,y}=\exp{(\Delta^{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta^{x,y}(E)$ is a copy of the coproduct acting on sites $x$ and $y$. This operator $S^{x,y}$ satisfies
\begin{equation}\label{symmetryH}
    \left[S_{x,y},H_{x,y}\right]=0
\end{equation}
i.e. $S_{x,y}$ is a symmetry of $H_{x,y}$. \\ Exploiting these considerations we write
\end{comment}
\begin{equation}
    \begin{split}
        H_{x,y}^{T}D&=(R_{x}R_{y})H_{x,y}(R_{x}R_{y})^{-1}\left(d_{x}\otimes\dd\right)\left(d_{y}\otimes\mathcal{D}_{\widehat{y}}\right)\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{\widehat{z}}\right)
        \\&=\left(R_{x}\exp{(E^{x})}\otimes \dd\right)\left(R_{y}\exp{(E^{y})}\otimes \mathcal{D}_{\widehat{y}}\right)H_{x,y}\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{\widehat{z}}\right)
        \\&=
        DH_{x,y}
    \end{split}
\end{equation}
 where we used \eqref{transpositionPropertyH} and \eqref{symmetryH} in the second equality. Thus, \eqref{edgeDualRealtion} is proved. 
 %\textbf{Reamrk}: it is important to notice that the elements of the duality matrix \eqref{dualityElements} are well defined only if at each site $x\in V$ and for every species $a\in \{2,\ldots,N\}$ the number of dual particles $r_{x}^{a}$ is lower or equal than the number of original particles $n_{a}^{x}$. This implies that the dynamics of the dual process is simpler, because described by a lower number of particles. By the way, it is possible to perform computations similar to the one made in this proof choosing instead of  \eqref{equationE} {\color{red} $E=\sum_{a=2}^{N}E_{1a}$ or $E=\sum_{a=2}^{N}E_{aa}$.}
 %{\color{blue} That's confusing. What is E?}

@@

\subsection{The process on a finite graph}
\paragraph{Action on the occupation variable of $\mathcal{L}_{x,y}^{rd}$}
\paragraph{Action on the occupation variable of $\mathcal{L}_{x}^{rd}$}
\subsection{The process on a chain}
\subsubsection{Proof of duality}
With this prove we answered the second and third questions stated at the beginning of this section. \\
\textbf{FROM HERE ON-WARDS STILL TO CHECK}\\

@@

@@

\appendix
\section{Oscillator realization}\label{appA}

 The basis element $ (E_{ab})_{a,b\in\{1,\ldots,N\}}$ of $gl(N)$ can also be thought as proper compositions of the rising, and lowering operators of th Heisenberg Lie algebra
  \begin{equation}\label{creationOperators}

@@

Therefore, we have that 
		E_{ab}=\oad_{a}\oa_{b}\qquad \forall a,b=1,\ldots N
fulfil the action on the basis vector \eqref{actionE} and the commutation relation \eqref{eq:comgl} 

@@

\begin{equation}
	\begin{split}
		H=\sum_{x,y\in \mathcal{E}}\omega_{x,y}H_{x,y}+\sum_{x\in V}\Gamma_{x}H_{x}
	\end{split}
\end{equation}
where the edge Hamiltonian is
\begin{equation}
H_{x,y}=\sum_{a,l=1}^{N}\left(e_{al}\otimes e_{la}-e_{ll}\otimes e_{aa}\right)
 \end{equation}
 and where the site Hamiltonian is
 \begin{equation}
H_{x}=\sum_{a,l=1}^{N}\alpha_{a}^{x}\left(e_{a,l}-e_{ll}\right)
\end{equation}
Moreover, by the \eqref{coproductC2} and by computing the explicit action of the diagonal part, we can write this Hamiltonian in function of the coproduct of the second Casimir
\begin{equation}
	H=\sum_{x,y\in \mathcal{E}}\omega_{x,y}\left\{\frac{1}{2}\Delta(C_{2})-2j(4j+N)\frac{1}{2}\mathbbm{1}\otimes\mathbbm{1}\right\}+\sum_{x\in V}\Gamma_{x}\sum_{a,l=1}^{N}\alpha_{a}^{x}\left(e_{a,l}-2j\mathbbm{1}\right)
\end{equation}
\end{comment}
\bibliography{ref}
\bibliographystyle{unsrt}

\end{document}
